---
title: "2017 Australian Open Data"
output: html_document
---

Load in the data
```{r}
library(tidyverse)
matches <- read_csv("https://raw.githubusercontent.com/JeffSackmann/tennis_slam_pointbypoint/master/2017-ausopen-matches.csv")
points <- read_csv("https://raw.githubusercontent.com/JeffSackmann/tennis_slam_pointbypoint/master/2017-ausopen-points.csv")
```

Isolate the matches that Roger Federer participated in.
```{r}
fed_pts <- matches %>%
  filter(grepl("Federer", player1) | grepl("Federer", player2)) %>%
  select(match_id, match_num, player1, player2) %>%
  left_join(points, by = "match_id") %>%
  arrange(match_num)
```

There are a lot of columns in these data.
Simplify things and focus in on just a few variables that I'll use in the plot.
```{r}
fp <- fed_pts %>%
  group_by(match_num) %>%
  mutate(point_duration = as.numeric(ElapsedTime - lag(ElapsedTime, default = 0))) %>%
  ungroup() %>%
  transmute(match_num, SetNo, GameNo,
            opponent = ifelse(player1 == "Roger Federer", player2, player1),
            play_time = cumsum(point_duration + 5) / 3600,
            serving = ifelse(player1 == "Roger Federer", PointServer == 1, PointServer == 2),
            won_point = ifelse(player1 == "Roger Federer", PointWinner == 1, PointWinner == 2)) %>%
  mutate(cum_points = cumsum(ifelse(won_point, 1, -1)),
         opponent = forcats::as_factor(opponent)
         )
```

Fit the GAM.
I could use `geom_smooth` here, but it is easier to control both the model and the transparency for the line and confidence interval separately by fitting the model myself.
```{r}
library(mgcv)
m <- gam(cum_points ~ s(play_time), data = fp)
pd <- tibble(play_time = seq(min(fp$play_time), max(fp$play_time), length.out = 300))
prediction <- predict(m, newdata = pd, se.fit = TRUE)
pd <- pd %>%
  mutate(cum_points = prediction$fit,
         upr = prediction$fit + qnorm(0.975) * prediction$se.fit,
         lwr = prediction$fit - qnorm(0.975) * prediction$se.fit)
```

Generate the plot.
Sorry about the complicated `theme` stuff.
I have all that bottled up into a package function, but that function is not publicly available anywhere.
```{r}
p <- ggplot(fp, aes(x = play_time, y = cum_points)) +
  geom_line(data = pd, color = "black", alpha = 2/3) +
  geom_ribbon(data = pd, aes(ymax = upr, ymin = lwr), alpha = 1/3) +
  geom_line(aes(color = opponent, alpha = as.numeric(serving))) +
  theme_mitre(legend_pos = "right", text_size = 12, legend_box = "vertical") +
  scale_alpha_continuous("Federer Serving",
                         range = c(1, 1/3),
                         breaks = c(1/3, 1),
                         labels = c("Yes", "No")) +
  scale_color_discrete("Opponent (and round)") +
  xlab("Approximate Play Time [hours]") +
  ylab("Points For - Points Against") +
  ggtitle("Roger Federer in the 2017 Australian Open") +
  theme_grey(base_size = 12,
             base_family = "Helvetica") +
  theme(axis.text = element_text(size = rel(0.8)),
        axis.ticks = element_line(colour = "black"), 
        strip.background = element_rect(fill = "grey90", 
                                        colour = "grey90"),
        legend.background = element_blank(), 
        legend.key = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        panel.grid.major.y = element_line(size = 0.7, 
                                          color = "grey", 
                                          linetype = "dashed"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        panel.background = element_rect(fill = NA, colour = NA), 
        plot.background = element_blank(),
        legend.position = "right", 
        legend.box = "vertical",
        legend.text.align = 1, 
        legend.title = element_text(size = rel(0.8)), 
        axis.title.x = element_text(size = rel(0.8), 
                                    margin = margin(t = 10, unit = "pt")),
        axis.title.y = element_text(size = rel(0.8), 
                                    margin = margin(r = 10, unit = "pt"),
                                    angle = 90), 
        text = element_text(size = 12,
                            family = "Helvetica", 
                            face = "plain", 
                            colour = "black",
                            hjust = 0.5, 
                            vjust = 0.5,
                            angle = 0,
                            lineheight = 1.1,
                            margin = margin(t = 5, b = 5),
                            debug = FALSE)) 
p
```

